A Switching Planner for Combined Task and Observation Planning

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
spelling
the value stuff clarrified
 - Charles

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
for each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\actions(p_i;(T_i))$ can be executed if no
$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
is in the root conjunct, or it occurs in $T_k$ for some executed
$\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^v$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^v$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

Finally, we come to valuing sequential plans, and therefore giving a
criterion that the sequential planner seeks to optimise. Where $p_i$
is the probability that the $i^{th}$ action execution is not a
transition to $s^\bot$, we value a trace according to:

@@

\noindent In other words, the value of a trace captures its
significance, in relation to other possible realisations of the plan,
to the overall expected value of the plan in
Eq~\ref{eq:expectedvalue}. In goal directed problems, where the only
non-zero reward is received at the first transition to a goal state,
Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
infamous degenerate hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy.

>>>>

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
for each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\actions(p_i;(T_i))$ can be executed if no
$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
is in the root conjunct, or it occurs in $T_k$ for some executed
$\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

Finally, we come to valuing sequential plans, and therefore giving a
criteria for the sequential planner to optimise. Where $p_i$ is the
probability that the $i^{th}$ sequenced action does not transition to
$s^\bot$, we value a trace as follows:

@@

\noindent The optimal trace given a plan, i.e., the non-assumptive
actions in sequence, therefore has value:

\[
V^* = \max_{p_1, .., p_n} \prod_{i=1..N} p_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]

\noindent which is equal to the maximal contribution a trace of the plan can
make to Eq~\ref{eq:expectedvalue}. In other words, writing $V_i$ for
the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
s_1, a_1,.., s_N$, we have:

\[
V^* = \arg\min_{V_i} \Expect \bigg{[} 
\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate = \bstate_0
\bigg{]} - V_i
\]

In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of the infamous degenerate
hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy.
